---
title: "hw3"
author: "Yue Chen"
date: "10/6/2020"
output: github_document
---

```{r}
library(tidyverse)
library(ggridges)
library(patchwork)
library(ggplot2)
library(scales)
library(hexbin)
library(magrittr)
```

## Problem 1
```{r}
library(p8105.datasets)
data("instacart")
```
The dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns. 
Observations are the level of items in orders by user. There are user / order variables - user ID, order ID, order day, and order hour. There are also item variables - name, aisle, department, and some numeric codes.

How many aisles are there, and which aisles are the most items ordered from?
```{r}
instacart %>%
  count(aisle) %>%
  arrange(desc(n))
```
There are 134 aisles. The most ordered aisles are fresh vegetables, fresh fruits, and packaged vegetable fruits.

Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
```{r}
instacart %>%
  count(aisle) %>%
  filter(n > 10000) %>%
  mutate(
    aisle = factor(aisle),
    aisle = fct_reorder(aisle, n)
  ) %>%
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.
```{r}
instacart %>%
  filter(aisle %in% c("baking ingredient", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>%
  count(product_name) %>%
  mutate(rank = min_rank(desc(n))) %>%
  filter(rank < 4) %>%
  arrange(aisle, rank) %>%
  knitr::kable()
```

Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers.
```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  ) %>%
  knitr::kable()

```

## Problem 2
```{r}
accel_data = read_csv("./accel_data.csv") %>%
  janitor::clean_names() %>%
  mutate(
    weekday = recode(day, 
                     "Monday" = "Weekday", 
                     "Tuesday" = "Weekday", 
                     "Wednesday" = "Weekday", 
                     "Thursday" = "Weekday", 
                     "Friday" = "Weekday", 
                     "Saturday" = "Weekend",
                     "Sunday" = "Weekend")) %>%
  pivot_longer(
    activity_1:activity_1440,
    names_to = "activity",
    names_prefix = "activity_", 
    values_to = "count"
  ) %>%
  mutate(
    week = as.integer(week),
    day_id = as.integer(day_id),
    activity = as.integer(activity),
    count = as.numeric(count)
  )

```

In the resulting dataset, there are `r nrow(accel_data)` observations and the following variables: `r names(accel_data)`. 

Using your tidied dataset, aggregate accross minutes to create a total activity variable for each day, and create a table showing these totals. 
```{r}
accel_data %>%
  group_by(day_id, day, weekday) %>%
  summarise(
    daily_count = sum(count)
  ) %>%
  knitr::kable()
```

It is hard to see a trend from the table.

Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. 

```{r}
accel_data %>%
  mutate(
    hour = activity / 60
  ) %>%
  group_by(hour, week, day, day_id) %>%
  summarise(activity_hour = sum(count)) %>%
  ggplot(aes(x = hour, y = activity_hour, color = day)) +
  geom_point(alpha = 0.6) +
  labs(
    title = "24-hour activity time courses for each day",
    x = "hour",
    y = "activity count in minutes"
  ) 
```
The patient is generally active at 6-7am on Thursdays, 11-12am on Sundays, 4-5pm on Saturdays, and 8-10pm on Fridays.The most vigorous activity happened on a Wednesday at 7-8pm.

## Problem 3
```{r}
library(p8105.datasets)
data("ny_noaa")
```

The dataset contains `r nrow(ny_noaa)` rows and `r ncol(ny_noaa)` columns. The variables are `r names(ny_noaa)`. 

There are `r sum(is.na(ny_noaa$snwd))` missing values in snow depth, accounting for `r percent(sum(is.na(ny_noaa$snwd)) / nrow(ny_noaa))` of all observations.

There are `r sum(is.na(ny_noaa$tmax))` missing values in maximum temperature, accounting for `r percent(sum(is.na(ny_noaa$tmax)) / nrow(ny_noaa))` of all observations.

There are `r sum(is.na(ny_noaa$tmin))` missing values in minimum temperature, accounting for `r percent(sum(is.na(ny_noaa$tmin)) / nrow(ny_noaa))` of all observations.

Data cleaning. Create separate variables for year, month, and day. Ensure observations for temperature, precipitation, and snowfall are given in reasonable units. 
```{r}
ny_noaa = 
ny_noaa %>%
  janitor::clean_names() %>%
  separate(col = date, into = c("year", "month", "day")) %>%
  mutate(
    year = as.integer(year),
    month = as.integer(month),
    day = as.integer(day),
    month = month.name[month],
    prcp = as.numeric(prcp, na.rm = T) / 10,
    tmax = as.numeric(tmax, na.rm = T) / 10,
    tmin = as.numeric(tmin, na.rm = T) / 10
  )
```
```{r}
ny_noaa %>%
  group_by(snow) %>%
  summarise(n = n()) %>%
  filter(min_rank(desc(n)) < 2)
```

The most observed value in snowfall is 0 because New York state has way less snow days than days without snow.

Make a two-panel plot showing the average max temperature in January and in July in each station across years.

```{r}
plot = 
ny_noaa %>% 
  filter(month%in% c("January", "July")) %>% 
  group_by(id, year, month) %>% 
  summarise(tmax_mean = mean(tmax)) %>% 
  ggplot(aes(x = year, y = tmax_mean, color = id, group = id)) +
  geom_point(alpha = 0.5) +
  geom_line(alpha = 0.2) +
  facet_grid(~ month) +
  theme(axis.text.x = element_text(angle = 45, hjust = 0.5, vjust = 0.5,size = 7),legend.position = "none")+
  labs(
    title = "Average max temperature in January and July in each station across years",
    x = "Year",
    y = "Average max temperature in Celsius"
  )

plot
```
In general, the average max temperature in January is lower than that in July. There is a greater fluctuation of average max temperature in January than that in July.

Some outliers are Jan 1982, Jul 1988, Jul 2004, and Jul 2007.

Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); and (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.

```{r}
plot1 = 
  ny_noaa %>%
  ggplot(
    aes(x = tmax, y = tmin)) + 
  geom_hex() + 
  geom_smooth() + 
  labs(
    title = "max vs min temperatures",
    x = "max temperature in Celsius", 
    y = "min temperature in Celcius", 
    caption = "Data from rnoaa package"
  ) 

plot2 = 
  ny_noaa %>%
  filter(snow > 0 & snow < 100) %>%
  ggplot(
    aes(x = snow, y = as.factor(year))
  ) + 
  geom_density_ridges() + 
  labs(
    title = "distribution of snowfall by year", 
    x = "snowfall in mm", 
    y = "year",
    caption = "Data from rnoaa package"
  )

two_panel_plot = plot1 + plot2

two_panel_plot
```


